Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem

### 2015-11-22 acta_manuscript.tex
added geo transfer results

\usepackage{acta_astronautica_packages}
\input{macros}

@@

\begin{keyword}
Keywords
\end{keyword}

@@

The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975}.
In this work, a multiple shooting method is implemented.
Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
Additional interior constraints are used to ensure a continuous solution.
This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
A numerical simulation is presented to demonstrate the transfer procedure.
The goal is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
The initial state is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the desired trajectories remain in the vicinity of the Moon in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.

@@

%Determine reachable set on \Poincare Section
%
%Intersection of the sets on section
%
%Minimization to a target state
%
%Show transfer to the moon
%
%Compare with the Ross work and the invariant manifold ( long time of flight)
\section{Conclusions}\label{sec:conclusion}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
This analysis has also assumed perfect attitude control and the ability to orient thrust in any direction.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability.

>>>>

\usepackage{acta_astronautica_packages} % packages and referencing customization
\input{macros} % defines custom macros

@@

\begin{keyword}
Low-thrust, Reachability, Optimal Control, Variational Integrator
\end{keyword}

@@

The optimal control formulation presented in this section results in a two point boundary value problem (TPBVP). 
There exist many methods to solve TPBVPs such as gradient, quasilinearization, and shooting methods~\cite{bryson1975}.
In this work, a multiple shooting method is implemented.
Shooting methods are common in astrodynamic trajectory design problems and relatively simple to implement.
In the shooting method, initial conditions are varied such that a terminal constraint is satisfied, similar to the way an archer modifies the bow in order to more accurately strike a target. 
Rather than numerical integration over the entire time interval, multiple shooting segments the interval into several smaller sub-arcs.
Additional interior constraints are used to ensure a continuous solution.
This has the benefit of decreasing the numerical sensitivity of the final states to changes in the initial conditions along each sub-arc.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
\subsection{Periodic Orbit transfer}
We present two numerical simulations to demonstrate the transfer procedure.
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix} \) with \( \mu = 0.0125 \).
The initial state is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{moon_orbit} % requires the graphicx package
   \caption{Target Orbit Region}
   \label{fig:moon_orbit}
\end{figure}
\Cref{fig:moon_orbit} shows that the desired trajectories remain in the vicinity of the Moon in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
For example, a series of communication satellites could be placed in this type of orbit. 
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.

@@

\subsection{Geostationary Transfer}
This example demonstrates the feasibility of iteratively computing the reachable set. 
In this manner, it is possible to design arbitrary transfers which require extended transfer durations.
Initially, it is assumed that the spacecraft lies on a circular geostationary orbit in the non-dimensional three body system. 
It is desired to transfer to a periodic orbit about the \( L_1 \) Lagrange point. 
The initial and final states of are illustrated in~\cref{fig:geo_transfer_target}.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.5\textwidth]{initial_final} % requires the graphicx package
   \caption{Transfer target}
   \label{fig:geo_transfer_target}
\end{figure}
In order to effectively utilize the dynamics of the three-body system the stable manifold of the periodic orbit is targeted.
This type of low energy transfer is best suited for autonomous spacecraft missions as the long time of flight is undesirable for manned spacecraft. 
Future proposals for permanent lunar spacecraft and bases will require frequent supply missions to remain viable. 
Low energy transfers from the Earth to the Lagrange points are necessary for future missions.

We generate the stable invariant manifold associated with the periodic orbit.
The stable manifold is propagated to the \Poincare section directed along the positive \( \hat{x} \) axis, or \( \alpha =\ang{0} \).
The reachable set is computed beginning on the geostationary orbit at it's intersection with the \Poincare section.
A final state is chosen from the reachable set to initialize subsequent iterations. 
The state which minimizes the distance on the \Poincare section to the stable manifold, illustrated in green in~\cref{fig:geo_transfer_poincare}, is used as the initial condition for the next reachable set computation.
With each reachable set we move the final trajectory closer to the stable manifold.
\Cref{fig:geo_transfer_poincare} shows that after eight iterations the resulting trajectory intersects the stable manifold. 
Combining these trajectories results in the powered portion of the transfer from the geostationary orbit to the stable manifold. 
\Cref{fig:geo_transfer_full,fig:geo_transfer_zoom} shows the resulting trajectory found, with the final trajectory shown which ensures the intersection with the stable invariant manifold.
Once at the stable manifold, no further control input is required and the vehicle will coast towards the target orbit.
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_full} 
		\caption{Transfer} \label{fig:geo_transfer_full} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{geo_transfer_zoom} 
		\caption{State Trajectory} \label{fig:geo_transfer_zoom} 
	\end{subfigure}

	\begin{subfigure}[htbp]{0.5\textwidth} 
		\includegraphics[width=\textwidth]{poincare} 
		\caption{\Poincare section} \label{fig:geo_transfer_poincare} 
	\end{subfigure} 
	\caption{Geostationary Transfer}
	\label{fig:geo_transfer} 
\end{figure}

\section{Conclusions}\label{sec:conclusion}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
This analysis has also assumed perfect attitude control and the ability to orient thrust in any direction.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability.
